---
title: "Physical Health"
output: html_notebook
---

# TBI

DATA -- ABCD Sum Scores Traumatic Brain Injury in `abcd_tbi01` 
  
Traumatic brain injury of youth summary scores (baseline) -- sums score from ABCD Parent Ohio State Traumatic Brain Injury Screen-Short Modified `abcd_otbi01`  
  
Variable of interest: `tbi_ss_worst_overall`  
Worst Injury Overall: Worst TBI -   
1, Improbable TBI (no TBI or TBI w/o LOC or memory loss);   
2, Possible mild TBI (TBI w/o LOC but memory loss);   
3, Mild TBI (TBI w/LOC ≤ 30 min);   
4, Moderate TBI (TBI w/LOC 30 min - 24 hrs);   
5, Severe TBI (TBI w/ LOC ≥ 24 hrs)  

citation: Bogner, J. A., Whiteneck, G. G., et al. (2017) Test-Retest Reliability of Traumatic Brain Injury Outcome Measures: A Traumatic Brain Injury Model Systems Study. J Head Trauma Rehabil 32(5): E1-E16.  

```{r}
rm(list=ls())
```


```{r}
data.dir <- ("~/Documents/abcd/physicalhealth/")
setwd(data.dir)
library(dplyr)

# read TBI
tbi = read.csv2("~/Documents/abcd/release2.0.1/abcd_tbi01.txt", sep = "\t", 
                header = TRUE, stringsAsFactors = F)
# delete the first row, which is the description of each column
# only include those without TBI and data from first assessment
notbi <- tbi[-c(1),] %>%
  filter(eventname == "baseline_year_1_arm_1", tbi_ss_worst_overall == 1) %>%
  arrange(subjectkey)
# then refresh row index
rownames(notbi) <- NULL

saveRDS(notbi, file = "notbi.rds")
notbi <- readRDS(file = "~/Documents/abcd/physicalhealth/notbi.rds")
```

# Family History Positive:  
   
Either Parent >= 2 symptoms: FHP 1108 
546F & 562M
     
Excluding TBI Either Parent >= 2 symptoms: FHP 1043   
518F & 525M
 

```{r}
fhp_thresh2 <- readRDS(file = "~/Documents/abcd/mentalhealth/fhp_thresh2.rds")

fhp2.notbi <- subset(fhp_thresh2, subjectkey %in% notbi$subjectkey)
fhp2.notbi <- select(fhp2.notbi, subjectkey, sex, interview_age)
  
saveRDS(fhp2.notbi, file = "~/Documents/abcd/mentalhealth/fhp_thresh2notbi.rds")
length(fhp2.notbi$subjectkey)
table(fhp2.notbi$sex)
```

Family History Negative: 6752         
3255F  &  3496M (1 unspecified)  
    
Excluding TBI: 6530    
3158  &  3371 (1 unspecified)  

```{r}
fhn <- readRDS(file = "~/Documents/abcd/mentalhealth/fhn.rds")

fhn.notbi <- subset(fhn, subjectkey %in% notbi$subjectkey)
fhn.notbi <- fhn.notbi %>% select(1:3)
  
saveRDS(fhn.notbi, file = "~/Documents/abcd/mentalhealth/fhnnotbi.rds")
length(fhn.notbi$subjectkey)
table(fhn.notbi$sex)
```


# Developmental History Questionnaire	(maternal substance use during pregnancy)

**Later can check with Fetal Alcohol Syndrome with IQ, head size, height and weight**

Raw data: ABCD Developmental History Questionnaire `dhx01` Kessler et al., 2009a, Kessler et al., 2009b, Merikangas et al., (2009)  

From Barch et al 2018  

Events and experiences that occurred either when the youth was in utero or throughout the course of their early development may either be indicators of early risk outcomes (delay in developmental milestones), or predictors of later outcomes (e.g., maternal substance use during pregnancy, birth trauma, etc.). Thus, the parent assessment battery includes a developmental history questionnaire that was originally developed by the Adolescent Component of the National Comorbidity Survey (Kessler et al., 2009a, Kessler et al., 2009b, Merikangas et al., 2009) with some additional supplemental questions on maternal use of substances during pregnancy (see Table 3). This questionnaire covers maternal prenatal care, maternal substance use during pregnancy (including caffeine and tobacco), prenatal maternal health conditions (e.g., gestational diabetes), prematurity, birth complications, and developmental milestones. This assessment is completed at the baseline initial assessment only.  

Summary data: ABCD Summary Scores Developmental History `abcd_devhxss01` -- summary

_ 8 _ variables: before knowing of pregnancy  
_ 9 _ variables: knowing of pregnancy  
1 = Yes /Sí;   exclude those
0 = No /No;   
999 = Don't know/ No lo sé    include those
devhx_ss_12_p = months premature birth

Tobacco?  How many times per day?  
Alcohol?  Maximum drinks in one sitting?  
Alcohol?  Average drinks per week?  
Alcohol?  How many drinks did it take to feel the effects of alcohol?  
Marijuana?   How many times per day?  
Cocaine/Crack? How many times per day?  
Heroin/Morphine?  How many times per day?  
Oxycontin?     How many times per day?  


Hill distinguish between FHAUD and alcohol exposure

```{r}
# read pregnancy drug exposure
preg = read.csv2("~/Documents/abcd/release2.0.1/abcd_devhxss01.txt", sep = "\t", 
                 header = TRUE, stringsAsFactors = F)
# delete the first row, which is the description of each column
# delete unneeded var, then EXCLUDE people who DIDNOT report NOT using cocain, heroin, oxycotin

preg <- preg[-c(1),] %>%
  filter(eventname == "baseline_year_1_arm_1") %>%
  arrange(subjectkey) %>%
  select(-devhx_ss_8_alcohol_max_p, -devhx_ss_8_alcohol_effects_p, 
         -devhx_ss_9_alcohol_max_p, -devhx_ss_9_alcohol_effects_p) %>%
  filter(devhx_ss_8_coc_crack_amt_p == 0, devhx_ss_8_her_morph_amt_p == 0,
           devhx_ss_8_oxycont_amt_p == 0, devhx_ss_9_coc_crack_amt_p == 0,
           devhx_ss_9_her_morph_amt_p == 0, devhx_ss_9_oxycont_amt_p == 0)

# then refresh row index
rownames(preg) <- NULL
preg[,10:22] <- sapply(preg[,10:22], as.numeric)

paste("devhx_ss_8_cigs_per_day_p")
table(preg$devhx_ss_8_cigs_per_day_p)
out1 <- mean(preg$devhx_ss_8_cigs_per_day_p, na.rm=T) + 2 * sd(preg$devhx_ss_8_cigs_per_day_p, na.rm=T)
paste("devhx_ss_8_alcohol_avg_p")
table(preg$devhx_ss_8_alcohol_avg_p)
out2 <- mean(preg$devhx_ss_8_alcohol_avg_p, na.rm=T) + 2 * sd(preg$devhx_ss_8_alcohol_avg_p, na.rm=T)
paste("devhx_ss_8_marijuana_amt_p")
table(preg$devhx_ss_8_marijuana_amt_p)
out3 <- mean(preg$devhx_ss_8_marijuana_amt_p, na.rm=T) + 2 * sd(preg$devhx_ss_8_marijuana_amt_p, na.rm=T)
paste("devhx_ss_9_cigs_per_day_p")
table(preg$devhx_ss_9_cigs_per_day_p)
out4 <- mean(preg$devhx_ss_9_cigs_per_day_p, na.rm=T) + 2 * sd(preg$devhx_ss_9_cigs_per_day_p, na.rm=T)
paste("devhx_ss_9_alcohol_avg_p")
table(preg$devhx_ss_9_alcohol_avg_p)
out5 <- mean(preg$devhx_ss_9_alcohol_avg_p, na.rm=T) + 2 * sd(preg$devhx_ss_9_alcohol_avg_p, na.rm=T)
paste("devhx_ss_9_marijuana_amt_p")
table(preg$devhx_ss_9_marijuana_amt_p)
out6 <- mean(preg$devhx_ss_9_marijuana_amt_p, na.rm=T) + 2 * sd(preg$devhx_ss_9_marijuana_amt_p, na.rm=T)

preg <- preg %>%
  filter(preg$devhx_ss_8_cigs_per_day_p < out1,
         preg$devhx_ss_8_alcohol_avg_p < out2,
         preg$devhx_ss_8_marijuana_amt_p < out3,
         preg$devhx_ss_9_cigs_per_day_p < out4,
         preg$devhx_ss_9_alcohol_avg_p < out5,
         preg$devhx_ss_9_marijuana_amt_p < out6)

saveRDS(preg, file = "preg.rds")
preg <- readRDS(file = "~/Documents/abcd/physicalhealth/preg.rds")
```
 
 
Family History Positive:  

Either Parent >= 2 symptoms: FHP 1108 
546F & 562M
     
Excluding TBI: FHP 1043   
518F & 525M

Excluding extreme drug exposure: FHP 605   
308F & 297M

```{r}
fhp2.notbi <- readRDS(file = "~/Documents/abcd/mentalhealth/fhp_thresh2notbi.rds")

fhp2.notbi.preg <- subset(fhp2.notbi, subjectkey %in% preg$subjectkey)
  
saveRDS(fhp2.notbi.preg, file = "~/Documents/abcd/mentalhealth/fhp_thresh2notbipreg.rds")
length(fhp2.notbi.preg$subjectkey)
table(fhp2.notbi.preg$sex)
```

Family History Negative: 6752         
3255F  &  3496M (1 unspecified)  
    
Excluding TBI: 6530    
3158  &  3371 (1 unspecified)  

Excluding drug exposure: 5538  
2679F  &  2858M (1 unspecified)

```{r}
fhn.notbi.preg <- subset(fhn.notbi, subjectkey %in% preg$subjectkey)
  
saveRDS(fhn.notbi.preg, file = "~/Documents/abcd/mentalhealth/fhnnotbipreg.rds")
length(fhn.notbi.preg$subjectkey)
table(fhn.notbi.preg$sex)
```

# other scales measuring physical health

Those were already analyzed in demographs

ABCD Sum Scores Physical Health Parent `abcd_ssphp01` -- different sleep disorders and puberty  
ABCD Sum Scores Physical Health Youth `abcd_ssphy01` -- puberty  
